---
title: "NATO Lane Change Test"
author: "Travis Baer"
date: "Tuesday, August 12, 2014"
output: html_document
---

Load data from database and run query to get data frame of one row per variant tested.

```{r results="hide",message=FALSE,warning=FALSE}
setwd("P:/External Projects/3104 - ALPS 2014/HMMWV Integration/Analysis")
require(RODBC)
require(ggplot2)
require(MASS)
require(DAAG)
require(car)
require(leaps)
require(randomForest)
require(gbm)
require(dplyr)
dataCN<-odbcConnectAccess("P:/External Projects/3104 - ALPS 2014/HMMWV Integration/Data/HMMWV Integration Master Data Consolidated")
#dataCN<-odbcConnectAccess("C:/Users/tbaer/Desktop/integration/HMMWV Integration Master Data Consolidated")
natoTest<-sqlQuery(dataCN,"SELECT * FROM NATOtest")
```

Clean up field names, switch the number of axles to a factor, remove the 4-axle test
```{r}
colnames(natoTest)<-gsub(" ","",colnames(natoTest))
colnames(natoTest)<-gsub("(","_",colnames(natoTest),fixed=TRUE)
colnames(natoTest)<-gsub("#","",colnames(natoTest),fixed=TRUE)
colnames(natoTest)<-gsub("/","_",colnames(natoTest),fixed=TRUE)
colnames(natoTest)<-gsub(")","_",colnames(natoTest),fixed=TRUE)
colnames(natoTest)<-gsub("-","",colnames(natoTest),fixed=TRUE)
colnames(natoTest)<-gsub(",","",colnames(natoTest),fixed=TRUE)
colnames(natoTest)<-gsub("%","Percent",colnames(natoTest),fixed=TRUE)
natoTest$Axles<-as.factor(natoTest$Axles)
sum(natoTest$Axles==4);natoTest<-natoTest[which(natoTest$Axles!=4),]
```


Plot with All Data
```{r}
p<-ggplot(natoTest,aes(x=TotalWeightAsTested,y=MaxSpeedNATO_mph_)) + geom_point(aes(shape=Axles,colour=WeightPercentHMMWVRearAxle,size=HMMWVVertCOG)) + labs(y="Max Speed",x="Total Weight")  + ggtitle("Max NATO")
plot(p)
hist(natoTest$MaxSpeedNATO_mph_,main="NATO Double Lane Change\nMaximum Entry Speed",xlab="mph") # exported at 400 width
plot(density(natoTest$MaxSpeedNATO_mph_))
# some mildly interesting things
plot(y=natoTest$MaxSpeedNATO_mph_,natoTest$WeightRatioHMMWVFront_Rear,xlab="HMMWV Weight Ratio: Front-Axle to Rear-Axle",
     main="Maximum Entry Speed Against\nHMMWV Longitudinal Weight Distribution",ylab="Maximum Entry Speed (mph)",pch=16,col="black")
points(y=natoTest$MaxSpeedNATO_mph_[18],x=natoTest$WeightRatioHMMWVFront_Rear[18],col="royalblue",pch=16) # export at 500 - # darkorange is alt color
abline(lm(natoTest$MaxSpeedNATO_mph_~natoTest$WeightRatioHMMWVFront_Rear),lty=2)
abline(lm(natoTest$MaxSpeedNATO_mph_[-18]~natoTest$WeightRatioHMMWVFront_Rear[-18]),lty=4)
natoTest[which(natoTest$MaxSpeedNATO_mph_==max(natoTest$MaxSpeedNATO_mph_)),"TestReportFilename"]
summary(lm(natoTest$MaxSpeedNATO_mph_~natoTest$WeightRatioHMMWVFront_Rear))
summary(lm(natoTest$MaxSpeedNATO_mph_[-18]~natoTest$WeightRatioHMMWVFront_Rear[-18]))
cor(natoTest$MaxSpeedNATO_mph_,natoTest$WeightRatioHMMWVFront_Rear)
cor(natoTest$MaxSpeedNATO_mph_[-18],natoTest$WeightRatioHMMWVFront_Rear[-18])
```

With Trailers
```{r}
p<-ggplot(natoTest,aes(x=TotalWeightAsTested,y=MaxSpeedNATO_mph_)) + geom_point(aes(shape=Axles,colour=WeightPercentHMMWVRearAxle,size=TrailerVertCOG)) + labs(y="Max Speed",x="Total Weight")  + ggtitle("Max NATO")
plot(p)
```

```{r, fig.height=10,fig.width=10}
natosub2<-natoTest[,c("TotalWeightAsTested","HMMWVVertCOG","WeightRatioHMMWVFront_Rear","HMMWVLatCOG","MaxSpeedNATO_mph_")]
spm(natosub2,smoother=FALSE)
natosub1<-natoTest[,c("TrailerLongCOG","WeightPercentComboTrailerAxle","TrailerVertCOG","TrailerLatCOG","MaxSpeedNATO_mph_")] # Limited to Trailers
spm(natosub1,smoother=FALSE)
natosub3<-natoTest[,c("Hlength","Hwidth","Hheight","MaxSpeedNATO_mph_")]
spm(natosub3,smoother=FALSE)
natosub4<-natoTest[,c("Tlength","Twidth","Theight","MaxSpeedNATO_mph_")]
spm(natosub4,smoother=FALSE)
# interesting ones
natosub5<-natoTest[,c("TotalWeightAsTested","Theight","WeightRatioHMMWVFront_Rear","MaxSpeedNATO_mph_")]
pairs(natosub5)
```

Which is a better fit: WeightRatioHMMWVFront_Rear or HMMWVLongCOG?
```{r}
aggregate(natoTest$MaxSpeedNATO_mph_,by=list("Axles"=natoTest$Axles),mean)
natoFitFR<-lm(MaxSpeedNATO_mph_ ~ WeightRatioHMMWVFront_Rear,data=natoTest)
natoFitHLongCG<-lm(MaxSpeedNATO_mph_ ~ HMMWVLongCOG,data=natoTest)
anova(natoFitFR,natoFitHLongCG) # all data
summary(natoFitFR)
plot(y=natoTest$MaxSpeedNATO_mph_,x=natoTest$WeightRatioHMMWVFront_Rear,col="orange",pch=16,
       ylab="Max Speed",xlab="HMMWV Front/Rear Weight Ratio")
abline(lm(natoTest$MaxSpeedNATO_mph_~natoTest$WeightRatioHMMWVFront_Rear),col="blue",lwd=2)
natoFitFR<-lm(MaxSpeedNATO_mph_ ~ WeightRatioHMMWVFront_Rear,data=natoTest[natoTest$Axles==2,])
natoFitHLongCG<-lm(MaxSpeedNATO_mph_ ~ HMMWVLongCOG,data=natoTest[natoTest$Axles==2,])
anova(natoFitFR,natoFitHLongCG) # HMMWV-only
plot(y=natoTest$MaxSpeedNATO_mph_[which(natoTest$Axles==2)],x=natoTest$WeightRatioHMMWVFront_Rear[which(natoTest$Axles==2)],col="orange",pch=16,
       ylab="Max Speed",xlab="HMMWV Front/Rear Weight Ratio")
abline(lm(natoTest$MaxSpeedNATO_mph_~natoTest$WeightRatioHMMWVFront_Rear,data=natoTest[natoTest$Axles==2,]),col="blue",lwd=2)
```

try a default RF, all data
```{r}
natoRightSpeedRF1st<-randomForest(MaxSpeedNATO_mph_~TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear,data=natoTest)
```

Stepwise Regression  - need to add in tire pressure soon - FULL MODEL
```{r}
#start<-lm(MaxSpeedNATO_mph_~TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+Hlength+Hwidth+Hheight,data=natoTest)
start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear)^2,data=natoTest)
start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear)^2+I(HMMWVLongCOG^2)+I(HMMWVLatCOG^2)+I(HMMWVVertCOG^2)+I(WeightRatioHMMWVFront_Rear^2)+I(log(HMMWVLongCOG))+I(log(HMMWVLatCOG))+I(log(HMMWVVertCOG))+I(log(WeightRatioHMMWVFront_Rear))+I(HMMWVLongCOG^2)+I(HMMWVLatCOG^1/2)+I(HMMWVVertCOG^1/2),data=natoTest)
#natoTopSpeedStepMany<-stepAIC(start)
natoTopSpeedLeaps<-regsubsets(formula(start),data=natoTest,nvmax=6,nbest=10,method="exhaustive")
#summary(natoTopSpeedLeaps)
plot(natoTopSpeedLeaps,scale="adjr2")
coef(natoTopSpeedLeaps,which(summary(natoTopSpeedLeaps)$adjr2==max(summary(natoTopSpeedLeaps)$adjr2))) # coefficients for the best model, still need to fit
natoTopSpeed1st<-lm(as.formula(paste("MaxSpeedNATO_mph_ ~",paste(names(coef(natoTopSpeedLeaps,which(summary(natoTopSpeedLeaps)$adjr2==max(summary(natoTopSpeedLeaps)$adjr2))))[-1],collapse="+"))) , data=natoTest)
summary(natoTopSpeed1st)
```

Diagnostics
```{r}
qqnorm(natoTopSpeed1st$residuals,main='Residual qqplot');qqline(natoTopSpeed1st$residuals)
natofitData1<-cbind(natoTopSpeed1st$model,residuals(natoTopSpeed1st))
spm(natofitData1,smoother=FALSE)
```


Stepwise Regression - Trailers only
```{r}
start<-lm(MaxSpeedNATO_mph_~TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+Hlength+Hwidth+Hheight+TrailerVertCOG+TrailerLatCOG+TrailerLongCOG+Tlength+Twidth+Theight+WeightPercentComboTrailerAxle,data=natoTest[natoTest$Axles==3,])
start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+Hlength+Hwidth+Hheight+TrailerVertCOG+TrailerLatCOG+TrailerLongCOG+Tlength+Twidth+WeightPercentComboTrailerAxle)^2,data=natoTest[natoTest$Axles==3,])
start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+TrailerVertCOG+TrailerLatCOG+TrailerLongCOG+WeightPercentComboTrailerAxle+WeightRatioComboHMMWV_Trailer+WeightRatioComboTongueWeight_TrailerAxleWeight+WeightAsTestedTotalTrailerWeight)^2+I(HMMWVLongCOG^2)+I(HMMWVLatCOG^2)+I(HMMWVVertCOG^2)+I(TrailerLongCOG^2)+I(TrailerLatCOG^2)+I(TrailerVertCOG^2)+I(WeightRatioHMMWVFront_Rear^2)+I(WeightPercentComboTrailerAxle^2)+I(log(HMMWVLongCOG))+I(log(HMMWVLatCOG))+I(log(HMMWVVertCOG))+I(HMMWVLongCOG^1/2)+I(HMMWVLatCOG^1/2)+I(HMMWVVertCOG^1/2)+I(HMMWVVertCOG^1/2)+I(TrailerLongCOG^1/2)+I(TrailerLatCOG^1/2)+I(WeightPercentComboTrailerAxle^1/2),data=natoTest[natoTest$Axles==3,]) # can't do logs b/c 0
#start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+TrailerVertCOG+TrailerLatCOG+TrailerLongCOG+WeightPercentComboTrailerAxle)^2+I(HMMWVLongCOG^1/2)+I(HMMWVLatCOG^1/2)+I(HMMWVVertCOG^1/2)+I(HMMWVVertCOG^1/2)+I(TrailerLongCOG^1/2)+I(TrailerLatCOG^1/2)+I(WeightPercentComboTrailerAxle^1/2)-HMMWVLongCOG:WeightRatioHMMWVFront_Rear,data=natoTest[natoTest$Axles==3,])
#start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+TrailerVertCOG+TrailerLatCOG+TrailerLongCOG+WeightPercentComboTrailerAxle)^2-HMMWVLongCOG:WeightRatioHMMWVFront_Rear+I(HMMWVLongCOG^2)+I(HMMWVLatCOG^2)+I(HMMWVVertCOG^2)+I(TrailerLongCOG^2)+I(TrailerLatCOG^2)+I(TrailerVertCOG^2)+I(WeightRatioHMMWVFront_Rear^2)+I(WeightPercentComboTrailerAxle^2),data=natoTest[natoTest$Axles==3,])
#natoTopSpeedTrailStepMany<-stepAIC(start)
natoTopSpeedTrailLeaps<-regsubsets(formula(start),data=natoTest[natoTest$Axles==3,],nbest=10,nvmax=6,method="exhaustive",really.big=T)
#summary(natoTopSpeedTrailLeaps)
plot(natoTopSpeedTrailLeaps,scale="adjr2")
plot(natoTopSpeedTrailLeaps,scale="bic")
coef(natoTopSpeedTrailLeaps,which(summary(natoTopSpeedTrailLeaps)$adjr2==max(summary(natoTopSpeedTrailLeaps)$adjr2))) # coefficients for the best model, still need to fit
natoTopSpeedTrail1st<-lm(as.formula(paste("MaxSpeedNATO_mph_ ~",paste(names(coef(natoTopSpeedTrailLeaps,which(summary(natoTopSpeedTrailLeaps)$adjr2==max(summary(natoTopSpeedTrailLeaps)$adjr2))))[-1],collapse="+"))) , data=natoTest)
natoTopSpeedTrail1stbic<-lm(as.formula(paste("MaxSpeedNATO_mph_ ~",paste(names(coef(natoTopSpeedTrailLeaps,which(summary(natoTopSpeedTrailLeaps)$bic==min(summary(natoTopSpeedTrailLeaps)$bic))))[-1],collapse="+"))) , data=natoTest) # returns same model
summary(natoTopSpeedTrail1st);summary(natoTopSpeedTrail1stbic)
plot(y=natoTopSpeedTrail1st$fitted.values,x=unlist(natoTopSpeedTrail1st$model[1]),pch=16,xlab="Actual Speed",ylab="Model-Predicted Speed",
     main="Actual vs. Predicted Entry Speed")
abline(0,1) # center line
points(x=natoTest$MaxSpeedNATO_mph_[18],y=max(natoTopSpeedTrail1st$fitted.values),pch=16,col="royalblue")
# confidence intervals for actual vs. predicted? I don't know how
newdata<-data.frame("TotalWeightAsTested"=16430,"HMMWVVertCOG"=39.4,"WeightRatioHMMWVFront_Rear"=0.6,"Theight"=66,"TrailerLatCOG"=0,"WeightPercentComboTrailerAxle"=0.2,"HMMWVLongCOG"=46.8,"TrailerLongCOG"=8,"TrailerVertCOG"=33.6,"Twidth"=44,"HMMWVLatCOG"=0.03,"Hheight"=72,TrailerVertCOG=35)
predict(natoTopSpeedTrail1st,newdata,interval="prediction",level=0.6)
toplotX<-data.frame("natoTest$MaxSpeedNATO_mph_"=seq(from=min(natoTest$MaxSpeedNATO_mph_),to=max(natoTest$MaxSpeedNATO_mph_),by=0.1))

# for comparison to previous model - drastically changed moodel
natoTopSpeedTrail1stadj<-lm(MaxSpeedNATO_mph_~TrailerVertCOG+WeightRatioHMMWVFront_Rear+HMMWVVertCOG:WeightRatioHMMWVFront_Rear+TrailerVertCOG:HMMWVLongCOG+WeightRatioHMMWVFront_Rear:TrailerLatCOG,data=natoTest) 
summary(natoTopSpeedTrail1stadj)
start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+TrailerVertCOG+TrailerLatCOG+TrailerLongCOG+WeightPercentComboTrailerAxle+WeightRatioComboHMMWV_Trailer+WeightRatioComboTongueWeight_TrailerAxleWeight+WeightAsTestedTotalTrailerWeight)^2+I(HMMWVLongCOG^2)+I(HMMWVLatCOG^2)+I(HMMWVVertCOG^2)+I(TrailerLongCOG^2)+I(TrailerLatCOG^2)+I(TrailerVertCOG^2)+I(WeightPercentComboTrailerAxle^2)+I(log(HMMWVLongCOG))+I(log(HMMWVVertCOG))+I(HMMWVLongCOG^1/2)+I(HMMWVLatCOG^1/2)+I(HMMWVVertCOG^1/2)+I(HMMWVVertCOG^1/2)+I(TrailerLongCOG^1/2)+I(TrailerLatCOG^1/2)+I(WeightPercentComboTrailerAxle^1/2),data=natoTest[natoTest$Axles==3,])# or an independent best subset
natoTopSpeedTrailLeaps<-regsubsets(formula(start),data=natoTest[natoTest$Axles==3,],nbest=10,nvmax=6,method="exhaustive",really.big=T)
plot(natoTopSpeedTrailLeaps,scale="adjr2")
natoTopSpeedTrail1stadj2<-lm(as.formula(paste("MaxSpeedNATO_mph_ ~",paste(names(coef(natoTopSpeedTrailLeaps,which(summary(natoTopSpeedTrailLeaps)$adjr2==max(summary(natoTopSpeedTrailLeaps)$adjr2))))[-1],collapse="+"))) , data=natoTest)
summary(natoTopSpeedTrail1stadj2)
newdf<-data.frame("TotalWeightAsTested"=13270.0,"HMMWVLatCOG"=1.00,"HMMWVLongCOG"=63.9,"HMMWVVertCOG"=37.2,"TrailerVertCOG"=46,"TrailerLongCOG"=9.6,
                  "WeightRatioHMMWVFront_Rear"=0.7,"TrailerLatCOG"=0.9,"WeightRatioComboHMMWV_Trailer"=2.6,WeightAsTestedTotalTrailerWeight=4107)
predict(natoTopSpeedTrail1stadj2,newdf,interval="prediction",level=0.90)
spm(natoTopSpeedTrail1stadj2$model)
```

```{r, results='hide',warning=FALSE}
dim(natoTopSpeedTrail1st$model)
CVSplits = 5#11#5 # how many groups to split the data into for cross validation?
fitWcv<-CVlm(df=natoTest[natoTest$Axles==3 & !is.na(natoTest$TrailerVertCOG),],form.lm=formula(natoTopSpeedTrail1st),m=CVSplits) ## Cross Validation
(cvR2<-1-sum(sum((fitWcv$MaxSpeedNATO_mph_-fitWcv$cvpred)^2))/sum((fitWcv$MaxSpeedNATO_mph_-mean(fitWcv$MaxSpeedNATO_mph_))^2))

dim(natoTopSpeedTrail1stadj2$model)
CVSplits = 6 # how many groups to split the data into for cross validation?
fitWcv<-CVlm(df=natoTest[natoTest$Axles==3,],form.lm=formula(natoTopSpeedTrail1stadj2),m=CVSplits) ## Cross Validation
(cvR2<-1-sum(sum((fitWcv$MaxSpeedNATO_mph_-fitWcv$cvpred)^2))/sum((fitWcv$MaxSpeedNATO_mph_-mean(fitWcv$MaxSpeedNATO_mph_))^2))
```

Diagnostics
```{r}
qqnorm(natoTopSpeedTrail1st$residuals,main='Residual qqplot');qqline(natoTopSpeedTrail1st$residuals)
natofitData1<-cbind(natoTopSpeedTrail1st$model,residuals(natoTopSpeedTrail1st))
spm(natofitData1,smoother=FALSE) # see residuals against predictors to see if a transform makes sense
```

```{r}
# predictions
newdf<-data.frame("TotalWeightAsTested"=13270.0,"HMMWVLatCOG"=1.00,"HMMWVLongCOG"=63.9,"HMMWVVertCOG"=37.2,"TrailerVertCOG"=46,
                  "WeightRatioHMMWVFront_Rear"=0.7,"TrailerLatCOG"=0.9,"WeightRatioComboHMMWV_Trailer"=2.6)
newdf<-data.frame("TotalWeightAsTested"=12060,"HMMWVLatCOG"=1.9,"HMMWVLongCOG"=46.8,"HMMWVVertCOG"=39.4,"TrailerVertCOG"=33.6,
                  "WeightRatioHMMWVFront_Rear"=0.5606641123882500,"TrailerLatCOG"=0)
predict(natoTopSpeedTrail1st,newdf,interval="prediction",level=0.90)
# calculate predictions with standard error estimtes
newdata<-newdf
res.var<-summary(natoTopSpeedTrail1st)$sigma^2; pred.var = res.var; Terms<-delete.response(terms(natoTopSpeedTrail1st))
m <- model.frame(Terms, newdata, na.action = na.pass,xlev = natoTopSpeedTrail1st$xlevels)
X <- model.matrix(Terms, m, contrasts.arg = natoTopSpeedTrail1st$contrasts) 
p<-natoTopSpeedTrail1st$rank;p1<-seq_len(p); piv<-natoTopSpeedTrail1st$qr$pivot[p1]
XRinv<-X[,piv] %*% qr.solve(qr.R(natoTopSpeedTrail1st$qr)[p1,p1]) # returns something the dimension of X
ip<-drop(XRinv^2 %*% rep(res.var,p))
beta<-natoTopSpeedTrail1st$coefficients
predictor<-sum(X*t(beta))#drop(X[,p,drop=FALSE] %*% beta[piv]) ## matrix multiplication isn't working? it doesn't add terms together
level1=0.80; df<-natoTopSpeedTrail1st$df.residual; tfrac<-qt((1-level1)/2,df,lower.tail=FALSE)
(hwid<-tfrac*sqrt(ip+pred.var))
(predictor + hwid %o% c(1,0,-1)) # prediction
predict(natoTopSpeedTrail1st,newdf,interval="prediction",level=level1)^2
write.csv(qr.solve(qr.R(natoTopSpeedTrail1st$qr)[p1,p1]),"Rinv_natoHMMWVTrailer.csv")
```

### Stepwise Regression - only HMMWVs
```{r}
start<-lm(MaxSpeedNATO_mph_~TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear+Hlength+Hwidth+Hheight,data=natoTest[natoTest$Axles==2,])
start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear)^2+I(HMMWVLongCOG^2)+I(HMMWVLatCOG^2)+I(HMMWVVertCOG^2)+I(WeightRatioHMMWVFront_Rear^2)+I(log(HMMWVLongCOG))+I(log(HMMWVLatCOG))+I(log(HMMWVVertCOG))+I(log(WeightRatioHMMWVFront_Rear))+I(HMMWVLongCOG^2)+I(HMMWVLatCOG^1/2)+I(HMMWVVertCOG^1/2),data=natoTest[natoTest$Axles==2,])
#start<-lm(MaxSpeedNATO_mph_~(TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+HMMWVLongCOG+WeightRatioHMMWVFront_Rear)^2+I(HMMWVLongCOG^2)-HMMWVLongCOG:WeightRatioHMMWVFront_Rear,data=natoTest[natoTest$Axles==2,])
#natoTopSpeedHMStepMany<-stepAIC(start)
natoTopSpeedHMLeaps<-regsubsets(formula(start),data=natoTest[natoTest$Axles==2,],nvmax=2,nbest=8,method="exhaustive")
# just take out length - I DID MESS UP - for final results need to use Longitidinal COG
#start<-lm(MaxSpeedNATO_mph_~TotalWeightAsTested+HMMWVVertCOG+HMMWVLatCOG+WeightRatioHMMWVFront_Rear+Hwidth+Hheight,data=natoTest[natoTest$Axles==2,])#summary(natoTopSpeedHMLeaps)
natoTopSpeedHMLeaps<-regsubsets(formula(start),data=natoTest[natoTest$Axles==2,],nvmax=2,nbest=8,method="exhaustive")
plot(natoTopSpeedHMLeaps,scale="adjr2")
coef(natoTopSpeedHMLeaps,which(summary(natoTopSpeedHMLeaps)$adjr2==max(summary(natoTopSpeedHMLeaps)$adjr2))) # coefficients for the best model, still need to fit
natoTopSpeedHM1st<-lm(as.formula(paste("MaxSpeedNATO_mph_ ~",paste(names(coef(natoTopSpeedHMLeaps,which(summary(natoTopSpeedHMLeaps)$adjr2==max(summary(natoTopSpeedHMLeaps)$adjr2))))[-1],collapse="+"))) , data=natoTest[natoTest$Axles==2,])
summary(natoTopSpeedHM1st)
plot(y=natoTopSpeedHM1st$fitted.values,x=unlist(natoTopSpeedHM1st$model[1]))
abline(0,1) # center line
predict(natoTopSpeedHM1st,newdata,interval="prediction",level=0.90)

natoTopSpeedHM2nd<-lm(MaxSpeedNATO_mph_~HMMWVVertCOG:HMMWVLatCOG + I(log(WeightRatioHMMWVFront_Rear)),data=natoTest[natoTest$Axles==2,])
summary(natoTopSpeedHM2nd)
```

```{r}
# predictions
newdf<-data.frame("TotalWeightAsTested"=12060,"HMMWVLatCOG"=1.00,"HMMWVLongCOG"=63.9,"HMMWVVertCOG"=37.2)
predict(natoTopSpeedHM1st,newdf,interval="prediction",level=0.90)
# calculate predictions with standard error estimtes
newdata<-newdf
res.var<-summary(natoTopSpeedHM1st)$sigma^2; pred.var = res.var; Terms<-delete.response(terms(natoTopSpeedHM1st))
m <- model.frame(Terms, newdata, na.action = na.pass,xlev = natoTopSpeedHM1st$xlevels)
X <- model.matrix(Terms, m, contrasts.arg = natoTopSpeedHM1st$contrasts) 
p<-natoTopSpeedHM1st$rank;p1<-seq_len(p); piv<-natoTopSpeedHM1st$qr$pivot[p1]
XRinv<-X[,piv] %*% qr.solve(qr.R(natoTopSpeedHM1st$qr)[p1,p1]) # returns something the dimension of X
ip<-drop(XRinv^2 %*% rep(res.var,p))
beta<-natoTopSpeedHM1st$coefficients
predictor<-sum(X*t(beta))#drop(X[,p,drop=FALSE] %*% beta[piv]) ## matrix multiplication isn't working? it doesn't add terms together
level1=0.80; df<-natoTopSpeedHM1st$df.residual; tfrac<-qt((1-level1)/2,df,lower.tail=FALSE)
(hwid<-tfrac*sqrt(ip+pred.var))
(predictor + hwid %o% c(1,0,-1)) # prediction
predict(natoTopSpeedHM1st,newdf,interval="prediction",level=level1)^2
write.csv(qr.solve(qr.R(natoTopSpeedHM1st$qr)[p1,p1]),"Rinv_natoHMMWV.csv")
spm(natoTopSpeedTrail1st$model,smooth=FALSE)
```

```{r, results='hide',warning=FALSE}
dim(natoTopSpeedHM1st$model)
CVSplits=3#6
fitWcv<-CVlm(df=natoTest[natoTest$Axles==2,],form.lm=formula(natoTopSpeedHM1st),m=CVSplits) ## Cross Validation
(cvR2<-1-sum(sum((fitWcv$MaxSpeedNATO_mph_-fitWcv$cvpred)^2))/sum((fitWcv$MaxSpeedNATO_mph_-mean(fitWcv$MaxSpeedNATO_mph_))^2))

fitWcv<-CVlm(df=natoTest[natoTest$Axles==2,],form.lm=formula(natoTopSpeedHM2nd),m=CVSplits) ## Cross Validation
(cvR2<-1-sum(sum((fitWcv$MaxSpeedNATO_mph_-fitWcv$cvpred)^2))/sum((fitWcv$MaxSpeedNATO_mph_-mean(fitWcv$MaxSpeedNATO_mph_))^2))
```

Diagnostics
```{r}
qqnorm(natoTopSpeedTrail1st$residuals,main='Residual qqplot');qqline(natoTopSpeedTrail1st$residuals)
natofitData1<-cbind(natoTopSpeedTrail1st$model,residuals(natoTopSpeedTrail1st))
spm(natofitData1,smoother=FALSE) # see residuals against predictors to see if a transform makes sense
```

